Abstractions and exemplars: the measure noun phrase alternation in German

### 2018-04-23 Latex/article.tex
Final spellchecking.

\begin{acknowledgement}
  I thank (in alphabetical order) Felix Bildhauer, Susanne Flach, Elizabeth Pankratz, Samuel Reichert, Ulrike Sayatz, and Christian Zimmer for valuable discussions and comments.
  Also, I would like to thank the reviewers for Cognitive Linguistics as well as associate editor Dagmar Divjak for insightful comments which helped to improve the quality of this paper significantly.
  Furthermore, I thank Ulrike Sayatz for helping me to recruit the participants for the experiments.
  Elizabeth Pankratz thankfully also fixed my English.
  Finally, I am grateful to my student assistants Kim Maser for her work on the annotation of the concordances and Luise Rißmann for supervising most of the experiments.
  The research presented here was made possible in part through funding from the \textit{Deutsche Forschungsgemeinschaft} (DFG, personal grant SCHA1916/1-1).
\end{acknowledgement}

>>>>

\begin{acknowledgement}
  I thank (in alphabetical order) Felix Bildhauer, Susanne Flach, Elizabeth Pankratz, Samuel Reichert, Ulrike Sayatz, and Christian Zimmer for valuable discussions and comments.
  Also, I would like to thank the three reviewers for Cognitive Linguistics as well as associate editor Dagmar Divjak for insightful comments which helped to improve the quality of this paper significantly.
  Furthermore, I thank Ulrike Sayatz for helping me to recruit the participants for the experiments.
  I am grateful to Sonya Faber and Elizabeth Pankratz (in alphabetical order) for fixing my English.
  Finally, I am grateful to my student assistants Kim Maser for her work on the annotation of the concordances and Luise Rißmann for supervising most of the experiments.
  The research presented here was made possible in part through funding from the \textit{Deutsche Forschungsgemeinschaft} (DFG, personal grant SCHA1916/1-1).
\end{acknowledgement}

### 2018-04-23 Latex/sections/1intro.tex
Final spellchecking.

This paper deals with a morpho-syntactic alternation between two measure noun constructions in German.
By \textit{alternation} I refer to a situation where two or more forms or constructions are available with no clear difference in acceptability, function, or meaning.
The study of lexical and constructional alternations has a long history in cognitive corpus linguistics (for example, \citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}).
This research is based on the assumption that language is a probabilistic phenomenon where alternants are chosen neither deterministically nor fully at random \citep{Bresnan2007}.
Multifactorial models are constructed which incorporate influencing factors from diverse levels, including lexical and contextual factors.
The estimation of the model coefficients quantifies the influence that the factors have on the probability that either alternant is chosen.
There are two fundamental issues to consider with respect to this tradition.
First, there is the question of whether corpus data provide any insight into cognitive representations at all.
This question can and should be answered by testing how well corpus-derived models converge with or diverge from experimental findings (Secion~\ref{sec:usagedataincognitivelinguistics}).
Second, the appropriate modelling of such results in cognitive linguistics is a key issue (Section~\ref{sec:prototypesexemplarscorporaandcontrolledexperiments}).

>>>>

This paper deals with a morpho-syntactic alternation between two measure noun constructions in German.
By \textit{alternation} I refer to a situation where two or more forms or constructions are available with no clear difference in acceptability, function, or meaning.
The study of lexical and constructional alternations has a long history in cognitive corpus linguistics (for example, \citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}).
This research is based on the assumption that language is a probabilistic phenomenon where alternants are chosen neither deterministically nor fully at random \citep{Bresnan2007}.
Multifactorial models are constructed which incorporate influencing factors from diverse levels, including lexical and contextual factors.
The estimation of the model coefficients quantifies the influence that the factors have on the probability that either alternant is chosen.
There are two fundamental issues to consider with respect to this tradition.
First, there is the question of whether corpus data provide any insight into cognitive representations at all.
This question can and should be answered by testing how well corpus-derived models converge with or diverge from experimental findings (Section~\ref{sec:usagedataincognitivelinguistics}).
Second, the appropriate modelling of such results in cognitive linguistics is a key issue (Section~\ref{sec:prototypesexemplarscorporaandcontrolledexperiments}).

### 2018-04-23 Latex/sections/4experiments.tex
Final spellchecking.

The marginal $R^2$ indicates a weak result for the fixed effects part of the model, which is nonetheless worthy of mention (close to 0.2).
The effect display for the single fixed regressor \textit{Modelprediction} is given in Figure \ref{fig:afc:effects}.
The higher the probability of the \PGCa\ predicted from usage data, the more often participants chose the \PGCa\ alternant in the forced-choice task.
A closer look at the results in the form of the spineplot in Figure~\ref{fig:spines} shows, however, that it was likely an idiosyncracy of a single sentence which spoiled an otherwise much better correlation.
The problematic sentence with a model prediction of 0.548 is given in (\ref{ex:problem}) in the \PGCa\ variant.

@@

\begin{figure}[htbp!]
  \centering
  \includegraphics[width=0.5\textwidth]{../R/output/fc_proportions}
  \caption{Spineplot of the proportion of responses plotted against the predictions from the corpus-based model in the forced-choice experiment}
  \label{fig:spines}
\end{figure}

In principle, random slope for \textit{Modelprediction} varying by item could also remedy problems with individual stimuli at least partially.
Therefore, a model with random slopes for \textit{Modelprediction} varying by both random effects (\textit{Participant} and \textit{Item}) was specified and estimated.
The random slope for participants was added to comply with \citet[257]{BarrEa2013} who predict ``catastrophically high Type I error rates'' for experimental designs with within-subject manipulations if random effects structures are not kept maximal.
The coefficient of the fixed effect changed noticeably but not enough to change the interpretation (5.408 relative to an intercept of -1.304), and the marginal $R^2_m$ rises to 0.213 ($R^2_c=0.488$).
In line with expectations, the standard deviation in the random slopes for \textit{Item} is high at 5.996.
However, the covariance parameters were estimated at exactly -1, which is a clear sign that the variance-covariance matrix could not be estimated successfully.
The same was true for models with only an \textit{Item} and a \textit{Participant} random slope.
This is exactly the kind of model overparametrisation criticised in \citet{BatesEa2015a} and \citet{MatuschekEa2017}.
The available data are insufficient to estimate the parameters of the more complex model with varying slopes.%
\footnote{\citet[1]{BatesEa2015a} state: ``We show that failure to converge typically is not due to a suboptimal estimation algorithm, but is a consequence of attempting to fit a model that is too complex to be properly supported by the data, irrespective of whether estimation is based on maximum likelihood or on Bayesian hierarchical modelling with uninformative or weakly informative priors.
Importantly, even under convergence, overparameterization may lead to uninterpretable models.''
}

>>>>

The marginal $R^2$ indicates a weak result for the fixed effects part of the model, which is nonetheless worthy of mention (close to 0.2).
The effect display for the single fixed regressor \textit{Modelprediction} is given in Figure \ref{fig:afc:effects}.
The higher the probability of the \PGCa\ predicted from usage data, the more often participants chose the \PGCa\ alternant in the forced-choice task.
A closer look at the results in the form of the spine plot in Figure~\ref{fig:spines} shows, however, that it was likely an idiosyncracy of a single sentence which spoiled an otherwise much better correlation.
The problematic sentence with a model prediction of 0.548 is given in (\ref{ex:problem}) in the \PGCa\ variant.

@@

\begin{figure}[htbp!]
  \centering
  \includegraphics[width=0.5\textwidth]{../R/output/fc_proportions}
  \caption{Spine plot of the proportion of responses plotted against the predictions from the corpus-based model in the forced-choice experiment}
  \label{fig:spines}
\end{figure}

In principle, random slope for \textit{Modelprediction} varying by item could also remedy problems with individual stimuli at least partially.
Therefore, a model with random slopes for \textit{Modelprediction} varying by both random effects (\textit{Participant} and \textit{Item}) was specified and estimated.
The random slope for participants was added to comply with \citet[257]{BarrEa2013} who predict ``catastrophically high Type I error rates'' for experimental designs with within-subject manipulations if random effects structures are not kept maximal.
The coefficient of the fixed effect changed noticeably but not enough to change the interpretation (5.408 relative to an intercept of -1.304), and the marginal $R^2_m$ rises to 0.213 ($R^2_c=0.488$).
In line with expectations, the standard deviation in the random slopes for \textit{Item} is high at 5.996.
However, the covariance parameters were estimated at exactly -1, which is a clear sign that the variance-covariance matrix could not be estimated successfully.
The same was true for models with only an \textit{Item} and a \textit{Participant} random slope.
This is exactly the kind of model overparametrisation criticised in \citet{BatesEa2015a} and \citet{MatuschekEa2017}.
The available data are insufficient to estimate the parameters of the more complex model with varying slopes.%
\footnote{\citet[1]{BatesEa2015a} state: ``We show that failure to converge typically is not due to a suboptimal estimation algorithm, but is a consequence of attempting to fit a model that is too complex to be properly supported by the data, irrespective of whether estimation is based on maximum likelihood or on Bayesian hierarchical modelling with uninformative or weakly informative priors.
Importantly, even under convergence, over-parameterisation may lead to uninterpretable models.''
}
